ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-07 rebuttal.tex

\begin{response}
    %
    Our intention is to describe the system to an audience that may not be familiar with the
    technical details of how the system is implemented. Thus, we feel that the length of the
    manuscript is a necessary outcome of providing the background for this information. However, the
    reviewer is correct that it is not ideal to leave much of the information in the documentation
    only. In combination with the suggestions of Reviewer 2, we have expanded the paper to provide
    more detail where necessary.
    %
\end{response}

\begin{reviewer}
    %
    p.~1 ``Competing standards such as the Cantera [2] CTI format or FlameMaster [3] lag behind
    considerably, although use of the former continually grows due to Cantera's open availability.''

    Note that FlameMaster is also freely available from the GitLab repository of the developers.
    %
\end{reviewer}

\begin{response}
    We attempted to access the GitLab repository at \url{https://git.rwth-aachen.de/ITV/FlameMaster}, but
    it is a closed repository. Thus, while it may be freely available to those with access, it is not
    \emph{openly} available.

    To clarify the meaning of ``open'' when used in our article, we have added a brief definition
    and link to the definition of open source, which comes from the Open Source Initiative.
\end{response}

@@

\begin{reviewer}
    %
    p.~2 ``Second, the XML format is intended to be a machine-readable markup language rather than a
    data format, and its lack of human readibility presents a barrier to creating and working with
    database files.''

    XML is a text format and it is fully human readable, similarly to YMAL. Even a
    first-time user is able to read the chemical content of a  PrIMe XML file and an experienced
    user is able to write such a file.
    %
\end{reviewer}

\begin{response}
    %
    We certainly recognize that XML is typically written in ASCII or UTF-8 (plain text) files, which
    a human can hypothetically decipher. Our argument is that XML is significantly harder for a
    human to read and write than YAML. This is primarily because XML has a lower signal-to-noise
    ratio than YAML; it requires many more characters to express a data structure in XML than in
    YAML. For instance, XML requires opening and closing tags, with the associated angle brackets,
    for every field while YAML does not; in XML, information can be specified as properties of tags
    or as elements (requiring the user to know where to look for information) while YAML has only
    three basic data structures (keys, values, and lists).

@@

    We note that although this verbosity in XML is a disadvantage when humans attempt to read or
    write such files, it allows much greater flexibility in specifying how a document should be
    marked up and provides XML with many advantages in other areas. However, since YAML is intended
    as a simple data format, and has a much simpler overall format, we argue that it is therefore
    much easier for humans to work with.

    We have modified the text on \textbf{page 2} to clarify this point, including describing the lack of human
    readability as a formatting issue (but one that is inherent to XML), rather than a technical
    issue.
    %
\end{response}

### 2017-10-07 pyked.tex
minor changes to close #23

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/}
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML is able to represent very detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be significantly more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.

>>>>

PrIMe~\autocite{Frenklach:2007bm,You:2011hy} (available at \url{http://primekinetics.org/}
has a number of features that make
it a strong standardization format, but it suffers from several flaws that
prevent wide adoption. First, the PrIMe standard does not require or support all
the information needed to simulate an experiment, including a machine-readable
definition of ignition or a standard way to express detailed facility-specific
effects necessary to properly simulate certain experiments. Second, the XML
format is intended to be a machine-readable markup language rather than a data format.
As a general-purpose document markup language, XML can represent detailed
information about a dataset, but implementing this detail requires XML-formatted
files to be much more verbose than files formatted with a language focused
only on representing data structures. The verbosity inherent in the XML format
limits the human readability of database files constructed in XML and presents a barrier to
creating and working with them. Furthermore, PrIMe uses
internal identifiers for bibliographic references and
species\slash reactions, rather than standard identifiers (e.g., DOIs for
scholarly products, InCHI or SMILES for species). This limits the usability of PrIMe files outside
the PrIMe ecosystem. Finally, the closed and opaque natures of the PrIMe standard
and associated Data Warehouse limit contributions and community development of
the PrIMe standard and compatible data files and tools.
